---
title: "Cell cycling analysis of bloodstream for trypanosoma brucei"
From: https://www.nature.com/articles/s41467-021-25607-2
---

```{r}
library(princurve)
library(tradeSeq)
library(cowplot)
library(clusterExperiment)
library(Seurat)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(rlang)
library(grid)
library(ggridges)
library(tidyverse)
library(slingshot)
library(ggridges)

# Require bar height map function for one heatmap, provided seperately 

# Plotting theme for graph asthetics
UMAP_theme <- theme(axis.line=element_blank(), axis.ticks = element_blank(),  panel.background = element_rect(size=0.5,linetype="solid",color="black"), plot.title = element_text(size = 10, face = "bold", hjust = 0.05, vjust = -8), axis.text.x = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y =  element_blank(), legend.title = element_blank())

```

```{r}
PrctCellExpringGene <- function(object, genes, group.by = "all"){
  if(group.by == "all"){
    prct = unlist(lapply(genes,calc_helper, object=object))
    result = data.frame(Markers = genes, Cell_proportion = prct)
    return(result)
  }
  
  else{        
    list = SplitObject(object, group.by)
    factors = names(list)
    
    results = lapply(list, PrctCellExpringGene, genes=genes)
    for(i in 1:length(factors)){
      results[[i]]$Feature = factors[i]
    }
    combined = do.call("rbind", results)
    return(combined)
  }
}

calc_helper <- function(object,genes){
  counts = object[['RNA']]@counts
  ncells = ncol(counts)
  if(genes %in% row.names(counts)){
    sum(counts[genes,]>0)/ncells
  }else{return(NA)}
}
```

### part one ###

```{r}
## Assign cell cycle stage using marker genes

Cell_cycle_regulated_genes <- read.delim("Cell_cycle_regulated_genes.txt")

## Can use the provided integrated seurat object or output from script "WT_differentiation_scRNA-seq"
load("Culture1")
# Rename object 
tryp <- mergeUMAP
DimPlot(tryp)
```


```{r}
# Get all the genes identified in the data set that are detected in at least 10% of the cells
genes <- tryp@assays[["RNA"]]@data
genes <- genes@Dimnames[[1]]
cell_prct <- PrctCellExpringGene(tryp, genes = genes, group.by = "all")
genes_10prct <- subset(cell_prct, subset = cell_prct$Cell_proportion > 0.1)
genes_10 <- genes_10prct$Markers

# Get list of marker genes present in at least 10% cells for each phase

Cell_cycle_regulated_genes <- read.delim("Cell_cycle_regulated_genes.txt")

s.genes <- subset(Cell_cycle_regulated_genes, Cell_cycle_regulated_genes$S.phase %in% genes_10)
s.genes <- s.genes$S.phase

g2m.genes <- subset(Cell_cycle_regulated_genes, Cell_cycle_regulated_genes$G2.M.phase %in% genes_10)
g2m.genes <- g2m.genes$G2.M.phase

early.g1.genes <- subset(Cell_cycle_regulated_genes, Cell_cycle_regulated_genes$Early.G1 %in% genes_10)
early.g1.genes <- early.g1.genes$Early.G1

late.g1.genes <- subset(Cell_cycle_regulated_genes, Cell_cycle_regulated_genes$Late.G1 %in% genes_10)
late.g1.genes <- late.g1.genes$Late.G1

```

```{r}
# Calculate an expression score for each phase and save it to the seurat object
tryp <- MetaFeature(tryp, features = s.genes, meta.name = "S.aggregate")
tryp <- MetaFeature(tryp, features = g2m.genes, meta.name = "G2M.aggregate")
tryp <- MetaFeature(tryp, features = early.g1.genes, meta.name = "Early.G1.aggregate")
tryp <- MetaFeature(tryp, features = late.g1.genes, meta.name = "Late.G1.aggregate")

# Creat and dataframe with the expression score of each cell and each phase
df <- data.frame(tryp@meta.data[["S.aggregate"]], tryp@meta.data[["G2M.aggregate"]], tryp@meta.data[["Early.G1.aggregate"]], tryp@meta.data[["Late.G1.aggregate"]])
colnames(df) <- c("S", "G2M", "Early G1", "Late G1")
rownames(df) <- tryp@assays[["RNA"]]@data@Dimnames[[2]]
# Find ratio between the score and average
df$S.ratio <- df$S / mean(df$S)
df$G2M.ratio <- df$G2M / mean(df$G2M)
df$Early_G1.ratio <- df$`Early G1` / mean(df$`Early G1`)
df$Late_G1.ratio <- df$`Late G1` / mean(df$`Late G1`)

# Find the top scoring phase of each cell, with FC > 1.5. 
assignments <- apply(
  X = df[, 5:8],
  MARGIN = 1,
  FUN = function(scores, first = 'S', second = 'G2M', third = "Early G1", fourth = "Late G1", null = 'Non-cycling') {
    if (all(scores < 1.5)) {
      return(null)
    } else {
      if (length(which(x = scores == max(scores))) > 1) {
        return('Undecided')
      } else {
        return(c(first, second, third, fourth)[which(x = scores == max(scores))])
      }
    }
  }    
)

df$Phase <- assignments
df$Cluster <- tryp@active.ident

head(df)

df$Phase <- assignments
df$Cluster <- tryp@active.ident

head(df)

#write.csv(df, file = "Cell_cycle_phase_scores.csv")

```

```{r}
mycolours <- c("#f8766d", "#7cae00", "#01bfc4", "grey", "#c77cff")
tryp$Phase <- as.vector(assignments)
DimPlot(object = tryp, group.by = "Phase", reduction = "umap", pt.size = 1, cols = mycolours) + UMAP_theme
p[[1]]$layers[[1]]$aes_params$alpha = 0.8
p[[1]]$layers[[1]]$aes_params$shape = 16
p

#pdf(file = "cycle_WT_umap.pdf", width = 3.7, height = 2.5)
#p
#dev.off()

```

```{r}

cell_proportions <- as.data.frame(prop.table(table(tryp$Phase, tryp@active.ident), margin = 2))
write.csv(cell_proportions, file = "cell_proportions.csv")

ggplot(data=cell_proportions, aes(x=cell_proportions$Var2, y=cell_proportions$Freq, fill=cell_proportions$Var1)) + geom_bar(stat="identity", color="black") + labs(x="sample", y="Proportion of Cells", fill="Cluster")


```




Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

